Approximation as Differentiation

### 2025-05-08 main/formal-structures.tex
Intro to Fam section.
Grothendieck.

>>>>

\subsection{A categorical framework for Galois slicing}

For a total language where types denote sets, the goal is to interpret programs as Galois connections relating
approximations of their input to approximations of their output. Since these lattice of approximations depends
on the value being approximated, the semantics of every type $\tau$ must have both a ``conventional''
component, assigning $\tau$ to a set $X$, and an associated lattice of approximations $\partial X(x)$ for
every point $x \in X$. The semantics of expression $e$ of type $Y$ with a single free variable of type $X$
must then also have two components: a function $f: X \to Y$ giving the baseline (unapproximated) meaning of
the expression, plus a family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for
every $x \in X$ which can be used for slicing over that expression.

This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
gloms all such $\Set$-indexed families together, with reindexing morphisms to transport between different
indexing sets.
